Should geophysicists use the gravity disturbance or the anomaly?

### 2018-05-22 manuscript/letter.tex
Final sentence for the abstract
Use new figures and place them in the text

\begin{summary}
 The gravity anomaly is defined as the difference between the Earth's gravity
 on the geoid and the gravity of the normal Earth on the reference ellipsoid.
 Because these quantities are not at the same point, the anomaly contains
 centrifugal accelerations and cannot be considered a harmonic function.
 The gravity disturbance is the difference between gravity and normal gravity
 at the same point.
 Consequently, the centrifugal effects can be neglected and the disturbance can
 be considered a harmonic function.
 This is the premise behind most potential-field data processing techniques
 (e.g., upward/downward continuation and vertical derivatives).
 Unlike the anomaly, the disturbance is due solely to the
 gravitational effects of geologic sources, making it the most appropriate
 for geophysical purposes.
 Use of the gravity anomaly in geophysics carries with it the implicit
 assumption that it is a good approximation for the gravity disturbance.
 However, bear in mind that the difference between the gravity disturbance and
 the free-air anomaly can be larger than 10 mGal worldwide.
 SOMETHING ABOUT MODELLING.
\end{summary}

@@

The first is a geocentric system of Cartesian coordinates
having the $Z$-axis coincident with the mean rotational axis of the Earth,
the $X$-axis pointing to the Greenwich meridian,
and the $Y$-axis completing a right-handed system (Fig. \ref{fig:GCS-GGS}).
This reference system is called by different names in the literature:
Mean Terrestrial System \citep[e.g.,][]{soler1976},
Earth-fixed geocentric Cartesian system \citep[e.g.,][]{torge2012}
or Earth-centered Earth-fixed system \citep[e.g.,][]{bouman_etal2013},
for example.
Here, we opted for using the term Geocentric Cartesian System (GCS).

The second is a geocentric system of geodetic coordinates
(the Geocentric Geodetic System or GGS),
which is defined by the reference ellipsoid used as the Normal Earth
\citep{heiskanen-moritz1967, soler1976, torge2012, bouman_etal2013}.
The position of a point is defined by
the \textit{geometric height} $h$,
the \textit{geodetic latitude} $\varphi$,
and the \textit{longitude} $\lambda$ (Fig. \ref{fig:GCS-GGS}).
There are also three mutually-orthogonal unit vectors defined at a given point
\citep{soler1976}:

@@

The third is the Topocentric Cartesian System (TCS),
which is commonly used in local or regional scale geophysical studies.
The origin of the TCS is located at a point $P$,
axes $x$ and $y$ are parallel to
the unit vectors $\versor{v}$ and $\versor{w}$
(eq. \ref{eq:unit-vectors}), respectively,
and the $z$-axis is opposite to the unit vector $\versor{u}$
(Fig. \ref{fig:TCS}).

@@

\noindent
where $\rho_t$ is the density of the topographic masses and $G$ is the
gravitational constant.
Th Bouguer correction approximates the gravitational effect that all
topographic masses above the geoid exert on point $P$ by the effect of a
homogeneous, infinitely extended slab of constant density $\rho_{t}$ and
thickness $H_{P}$.
A similar correction can be applied to the gravity disturbance
(using the approximation in eq. \ref{eq:gravity-disturbance-approx})
to remove the effect of topographic masses between
the observation point $P$ and a point on the ellipsoid $Q$.
We will call this quantity the \textit{Bouguer disturbance}:

@@

At first glance, the only difference between eqs.
\ref{eq:simple-bouguer-anomaly} and \ref{eq:bouguer-disturbance} is the use of
orthometric ($H_P$) versus geometric ($h_P$) heights.
However, their geophysical interpretations are significantly different.
Apart from the centrifugal effect mentioned in the previous section,
the Bouguer anomaly (eq. \ref{eq:simple-bouguer-anomaly}) still contains
the gravitational effect of topographic masses between the geoid and ellipsoid.
Conversely, when calculating the Bouguer disturbance one removes the effects of
an ellipsoidal reference model (the normal Earth) and the known topographic
masses between the surface of the ellipsoid and observation point.
Hence, the Bouguer disturbance can be interpreted as being solely the
gravitational effect produced by the density contrasts between the actual
internal density distribution of the Earth and that of the normal Earth plus a
homogeneous topography.

@@

The gravity vector $\vector{g}_i$ (eq. \ref{eq:gravity-disturbance-vector})
at a point $(x_i, y_i, z_i)$ in the TCS (Fig. \ref{fig:TCS}a) can be
represented by:

@@

\noindent
where $^{\top}$ denotes transposition and $\versor{\gamma}_{i}$ is a unit
vector with the direction of the normal gravity vector.
This approximation is valid because
$\gamma_{i} \gg \| \delta \mathbf{g}_{i} \|$ at all points on or above the
Earth's surface.
It is known in geodesy \citep[e.g.,][]{sanso_sideris2013}
and is analogous to the approximation used for
total-field magnetic anomalies \citep[e.g.,][]{blakely1996}.
In local or regional scales, the unit
vector $\versor{\gamma}_{i}$
may be considered constant throughout the study area and
parallel to the $z$ axis of the TCS (Fig. \ref{fig:TCS}a).

Using eq. \ref{eq:gobs-approx} and fixing $\gamma$ at a point $P$ at
the origin of the TCS (Fig. \ref{fig:TCS}a),
the gravity disturbance can be rewritten as:

@@

\noindent
This equation shows that the gravity disturbance $\delta g_{i}$ is
approximately the component of the $\delta \vector{g}_{i}$ in the direction of
the normal gravity vector $\vector{\gamma}_P$
\citep{hofmann-wellenhof-moritz2005, sanso_sideris2013}.
Because $z$-axis of the TCS is parallel to $\vector{\gamma}_P$
(Fig. \ref{fig:TCS}b), the gravity disturbance can be defined as the
$z$-component of the gravitational attraction exerted by the gravity sources at
the point $(x_{i}, y_{i}, z_{i})$.
As a consequence, the gravity disturbance can be modelled by the harmonic
function:

@@

In practice, most instances of gravity modelling in the literature use Bouguer
anomalies and function $d_{i}$ (usually presented as ``$g_z$'') to represent
the gravitational effect of gravity sources.
Consequently, these studies are using the gravity
anomaly as an approximation for the gravity disturbance.
Notice that $d_{i}$ does not depend on the
height with respect to the geoid (orthometric height).
Rather, it depends on the relative position of the gravity sources and the
observation point $(x_{i}, y_{i}, z_{i})$ in the TCS (Fig. \ref{fig:TCS}a).
Finally, it is important to mention that gravimeters do not measure $d_{i}$ but
$\| \vector{g}_i \|$.

@@

\appendix
%\section{For authors}
%
%Table~\ref{authors} is a list of design macros which are unique to GJI. The
%list displays each macro's name and description.
%
%\section{For editors}
%
%The additional features shown in Table~\ref{editors} may be used for
%production purposes.
%
%\bsp % ``This paper has been produced using the Blackwell
%     %   Publishing GJI \LaTeXe\ class file.''

\begin{figure}
%    \vspace{5.5cm}
    \includegraphics{figures/cartesian-geodetic-systems.png}
    \caption{
    The Geocentric Cartesian System (GCS) and the Geocentric Geodetic System
    (GGS).
    The GGS is defined by an oblate ellipsoid with semi-minor axis $b$ and
    semi-major axis $a$.
    In this coordinate system, the position of a point is determined by the
    geometric height $h$, geodetic latitude $\varphi$, and longitude $\lambda$.
    $O$ is the Earth's center of mass, $P$ is an observation point, and
    unit vectors $\versor{u}$, $\versor{v}$, and $\versor{w}$ define mutually
    orthogonal directions at $P$ (eq. \ref{eq:unit-vectors}).
    In (b), $Q$ is the projection of $P$ onto the reference ellipsoid at
    the same latitude and longitude.
    }
  \label{fig:GCS-GGS}
\end{figure}

\begin{figure}
    \includegraphics{figures/local-system.png}
    \caption{
    (a) The topocentric Cartesian coordinate system (TCS) with origin at a
    point $P$.
    Axes $x$ and $y$ are parallel to the unit vectors $\versor{v}$ and
    $\versor{w}$ (eq. \ref{eq:unit-vectors} and Fig. \ref{fig:GCS-GGS}),
    respectively, and the $z$ axis is opposite to $\versor{u}$.
    The gray plane is the same shown in Fig. \ref{fig:GCS-GGS}.
    (b) Schematic representation of the gravity vector
    $\vector{g}_{P}$, normal gravity vector $\vector{\gamma}_{P}$,
    gravity disturbance vector $\delta \vector{g}_{P}$
    (eq. \ref{eq:gravity-disturbance-vector}) and unit vector
    $\versor{u}_{P}$ (eq. \ref{eq:unit-vectors}) at a point
    $P$.
    }
  \label{fig:TCS}
\end{figure}

\label{lastpage}

>>>>

\begin{summary}
 The gravity anomaly is defined as the difference between the Earth's gravity
 on the geoid and the normal gravity on the reference ellipsoid.
 Because these quantities are not at the same point, the anomaly contains
 centrifugal accelerations and cannot be considered a harmonic function.
 The gravity disturbance is the difference between gravity and normal gravity
 at the same point.
 Consequently, the centrifugal effects can be neglected and the disturbance can
 be considered a harmonic function.
 This is the premise behind most potential-field data processing techniques
 (e.g., upward/downward continuation).
 Unlike the anomaly, the disturbance is due solely to the
 gravitational effects of geologic sources, making it the most appropriate
 for geophysical purposes.
 Use of the gravity anomaly in geophysics carries with it the implicit
 assumption that it is a good approximation for the gravity disturbance.
 However, bear in mind that the difference between the gravity disturbance and
 the free-air anomaly can be larger than 10 mGal worldwide.
 In fact, we argue that the assumptions made during gravity forward and inverse
 modeling imply that the quantity being modelled is the disturbance, not the
 anomaly.
\end{summary}

@@

The first is a geocentric system of Cartesian coordinates
having the $Z$-axis coincident with the mean rotational axis of the Earth,
the $X$-axis pointing to the Greenwich meridian,
and the $Y$-axis completing a right-handed system (Fig. \ref{fig:systems}).
This reference system is called by different names in the literature:
Mean Terrestrial System \citep[e.g.,][]{soler1976},
Earth-fixed geocentric Cartesian system \citep[e.g.,][]{torge2012}
or Earth-centered Earth-fixed system \citep[e.g.,][]{bouman_etal2013},
for example.
Here, we opted for using the term Geocentric Cartesian System (GCS).

The second is a geocentric system of geodetic coordinates
(the Geocentric Geodetic System or GGS),
which is defined by the reference ellipsoid used as the Normal Earth
\citep{heiskanen-moritz1967, soler1976, torge2012, bouman_etal2013}.
The position of a point is defined by
the \textit{geometric height} $h$,
the \textit{geodetic latitude} $\varphi$,
and the \textit{longitude} $\lambda$ (Fig. \ref{fig:systems}).
There are also three mutually-orthogonal unit vectors defined at a given point
\citep{soler1976}:

@@

The third is the Topocentric Cartesian System (TCS),
which is commonly used in local or regional scale geophysical studies.
The origin of the TCS is located at a point $P$,
axes $x$ and $y$ are parallel to
the unit vectors $\versor{v}$ and $\versor{w}$
(eq. \ref{eq:unit-vectors}), respectively,
and the $z$-axis is opposite to the unit vector $\versor{u}$
(Fig. \ref{fig:systems}c).

\begin{figure}
%    \vspace{5.5cm}
    \includegraphics{figures/cartesian-geodetic-systems.eps}
    \caption{
    (a) and (b) The Geocentric Cartesian System (GCS) and the Geocentric
    Geodetic System (GGS).
    The GGS is defined by an oblate ellipsoid with semi-minor axis $b$ and
    semi-major axis $a$.
    In this coordinate system, the position of a point is determined by the
    geometric height $h$, geodetic latitude $\varphi$, and longitude $\lambda$.
    $O$ is the Earth's center of mass, $P$ is an observation point, and
    unit vectors $\versor{u}$, $\versor{v}$, and $\versor{w}$ define mutually
    orthogonal directions at $P$ (eq. \ref{eq:unit-vectors}).
    In (b), $Q$ is the projection of $P$ onto the reference ellipsoid at
    the same latitude and longitude.
    (c) The topocentric Cartesian coordinate system (TCS) with origin at a
    point $P$.
    Axes $x$ and $y$ are parallel to the unit vectors $\versor{v}$ and
    $\versor{w}$, respectively, and the $z$ axis is opposite to $\versor{u}$.
    The gray plane is the same shown in (a) and (b).
    }
  \label{fig:systems}
\end{figure}

@@

\noindent
where $\rho_t$ is the density of the topographic masses and $G$ is the
gravitational constant.
The Bouguer correction approximates the gravitational effect that all
topographic masses above the geoid exert on point $P$ by the effect of a
homogeneous, infinitely extended slab of constant density $\rho_{t}$ and
thickness $H_{P}$.
A similar correction can be applied to the gravity disturbance
(using the approximation in eq. \ref{eq:gravity-disturbance-approx})
to remove the effect of topographic masses between
the observation point $P$ and a point on the ellipsoid $Q$.
We will call this quantity the \textit{Bouguer disturbance}:

@@

At first glance, the only difference between eqs.
\ref{eq:simple-bouguer-anomaly} and \ref{eq:bouguer-disturbance} is the use of
orthometric ($H_P$) versus geometric ($h_P$) heights.
However, their geophysical interpretations are significantly different.
Apart from the centrifugal effect mentioned in the previous section,
the Bouguer anomaly (eq. \ref{eq:simple-bouguer-anomaly}) still contains
the gravitational effect of topographic masses between the geoid and ellipsoid.
Conversely, when calculating the Bouguer disturbance one removes the effects of
an ellipsoidal reference model (the normal Earth) and the known topographic
masses between the surface of the ellipsoid and observation point.

@@

The gravity vector $\vector{g}_i$ (eq. \ref{eq:gravity-disturbance-vector})
at a point $(x_i, y_i, z_i)$ in the TCS (Fig. \ref{fig:systems}c) can be
represented by:

@@

\noindent
where $^{\top}$ denotes transposition and $\versor{\gamma}_{i}$ is a unit
vector with the direction of the normal gravity vector.
This approximation is valid because
$\gamma_{i} \gg \| \delta \mathbf{g}_{i} \|$ at all points on or above the
Earth's surface.
It is known in geodesy \citep[e.g.,][]{sanso_sideris2013}
and is analogous to the approximation used for
total-field magnetic anomalies \citep[e.g.,][]{blakely1996}.
In local or regional scales, the unit
vector $\versor{\gamma}_{i}$
may be considered constant throughout the study area and
parallel to the $z$ axis of the TCS (Fig. \ref{fig:systems}c).

Using eq. \ref{eq:gobs-approx} and fixing $\gamma$ at a point $P$ at
the origin of the TCS (Fig. \ref{fig:systems}c),
the gravity disturbance can be rewritten as:

@@

\noindent
This equation shows that the gravity disturbance $\delta g_{i}$ is
approximately the component of the $\delta \vector{g}_{i}$ in the direction of
the normal gravity vector $\vector{\gamma}_P$
\citep{hofmann-wellenhof-moritz2005, sanso_sideris2013}.
Because $z$-axis of the TCS is parallel to $\vector{\gamma}_P$
(Fig. \ref{fig:surfaces}), the gravity disturbance can be defined as the
$z$-component of the gravitational attraction exerted by the gravity sources at
the point $(x_{i}, y_{i}, z_{i})$.
As a consequence, the gravity disturbance can be modelled by the harmonic
function:

@@

In practice, most instances of gravity modelling in the literature use Bouguer
anomalies and function $d_{i}$ (usually presented as ``$g_z$'') to represent
the gravitational effect of gravity sources.
Consequently, these studies are using the gravity
anomaly as an approximation for the gravity disturbance.
Notice that $d_{i}$ does not depend on the
height with respect to the geoid (orthometric height).
Rather, it depends on the relative position of the gravity sources and the
observation point $(x_{i}, y_{i}, z_{i})$ in the TCS (Fig. \ref{fig:systems}c).
Finally, it is important to mention that gravimeters do not measure $d_{i}$ but
$\| \vector{g}_i \|$.

@@
